Efficient evaluation of threshold queries of derived fields in a numerical simulation database

### 2014-10-20 paper.tex
Added cache diagram.

\begin{figure}
\centering
\includegraphics[width=3.35in]{Figures/jhtdb_diagram.pdf}
\caption{Architectrue of the JHTDB.}
\label{fig:jhtdb_diagram}
\end{figure}

@@

%\begin{figure*}
%\centering
%\includegraphics[width=6.5in]{Figures/cache_diagram.pdf}
%\caption{Cache lookup.}
%\label{fig:cache_diagram}
%\end{figure*}

@@

{\bf Application-aware cache for query results:} A central part of the evaluation strategy for threshold queries that we have developed is the 
introduction of an application-aware cache
for query results. The results of these queries are small compared to the amount of data that need to be examined and the results can be used to
answer subsequent queries as long as they are within the same region and specify the same or higher threshold. Each database node has a local 
cache. Cache entries are indexed by the field, time-step, spatial region and the threshold requested. We use a least recently
used cache replacement policy. All modifications of and queries to the cache are executed within a transaction with snapshot isolation level to
avoid dirty-reads or an inconsistent view of the cache.

@@

It is interesting to note that the time taken to perform a cache lookup is relatively small even in the case of a cache hit as can be seen in Figures 
\ref{fig:vorticity_breakdown_cachehit}, \ref{fig:q_breakdown_cachehit}, \ref{fig:magnetic_breakdown_cachehit}. This is because the cache tables reside on
SSDs attached to each database node and retrieving the data is always done through a clustered index lookup. In the cases of a cache hit the majority of
the time is spent simply transferring the results from the database nodes to the mediator and then back to the user. These times remain more or less
constant between the cases of cache misses and cache hits (top row and bottom row of Fig. \ref{fig:breakdown}).
Caching the results of threshold queries
effectively preserves the I/O and computational effort spent during their initial evaluation and results in over an order of magnitude speedup for all the
different fields requested as we can see in Figure \ref{fig:breakdown}.

@@

Andrade et al. \cite{Andrade} describe a database system and an optimization framework build on the concept of \emph{active semantic caching}. An
active semantic cache aims to fully or partially reuse cached query results or aggregates through automated transformations of these aggregates. Similarly
to our work they focus on real scientific data-analysis applications. The method that we have developed for the evaluation of threshold queries complements
the active semantic caching approach and could be used in that framework. Our work has focused on extending a relational database system (Microsoft's SQL
Server) as opposed to designing a new database system from the ground up as Andrade et al. \cite{Andrade}.

>>>>

\begin{figure}
\centering
\includegraphics[width=3.35in]{Figures/jhtdb_diagram.pdf}
\caption{Architecture of the JHTDB.}
\label{fig:jhtdb_diagram}
\end{figure}

@@

@@

{\bf Application-aware cache for query results:} A central part of the evaluation strategy for threshold queries that we have developed is the 
introduction of an application-aware cache
for query results (Fig. \ref{fig:cache_diagram}). The results of these queries are small compared to the amount of data that need to be examined and the
results can be used to answer subsequent queries as long as they are within the same region and specify the same or higher threshold. Each database node 
has a local cache. Cache entries are indexed by the field, time-step, spatial region and the threshold requested. We use a least recently
used cache replacement policy. All modifications of and queries to the cache are executed within a transaction with snapshot isolation level to
avoid dirty-reads or an inconsistent view of the cache.

\begin{figure}
\centering
\includegraphics[width=3.35in]{Figures/cache_diagram.pdf}
\caption{Distributed evaluation of threshold queries and architecture of the application-aware cache. 
Each database node has local cache tables, which reside on solid-state drives attached to the node.}
\label{fig:cache_diagram}
\end{figure}

@@

It is interesting to note that the time taken to perform a cache lookup is relatively small even in the case of a cache hit as can be seen in Figures 
\ref{fig:vorticity_breakdown_cachehit}, \ref{fig:q_breakdown_cachehit}, \ref{fig:magnetic_breakdown_cachehit}. This is because the cache tables reside on
SSDs attached to each database node (see Fig. \ref{fig:cache_diagram}) and retrieving the data is always done through a clustered index lookup.
In the cases of a cache hit the majority of the time is spent simply transferring the results from the database nodes to the mediator and then back to the user.
These times remain more or less constant between the cases of cache misses and cache hits (top row and bottom row of Fig. \ref{fig:breakdown}).
Caching the results of threshold queries
effectively preserves the I/O and computational effort spent during their initial evaluation and results in over an order of magnitude speedup for all the
different fields requested as we can see in Figure \ref{fig:breakdown}.

@@

Andrade et al. \cite{Andrade} describe a database system and an optimization framework build on the concept of \emph{active semantic caching}. An
active semantic cache aims to fully or partially reuse cached query results or aggregates through automated transformations of these aggregates. Similarly
to our work they focus on real scientific data-analysis applications. The method that we have developed for the evaluation of threshold queries complements
the active semantic caching approach and could be used in that framework. Our work has focused on extending a relational database system (Microsoft's SQL
Server) as opposed to designing a new database system from the ground up as described by Andrade et al. \cite{Andrade}.
